---
title: "barY Mark"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mosaic)
```

The `barY` mark creates vertical bar charts. It draws rectangular bars from a baseline (typically zero) to the specified y value.

## Basic Example

```{r}
# Sample data
data <- data.frame(
  category = c("A", "B", "C", "D", "E", "F", "G", "H"),
  value = c(2, 8, 3, 7, 5, 4, 6, 1)
)

spec <- list(
  plot = list(
    list(
      mark = "barY",
      data = list(from = "sample_data"),
      x = "category",
      y = "value",
      fill = "steelblue"
    )
  )
)

mosaic(spec, sample_data = data)
```

## Customization Options

### Color by Category

```{r}
spec_colored <- list(
  plot = list(
    list(
      mark = "barY",
      data = list(from = "sample_data"),
      x = "category",
      y = "value",
      fill = list(column = "category")
    )
  )
)

mosaic(spec_colored, sample_data = data)
```

### Grouped Bars

```{r}
# Sample data with groups
grouped_data <- data.frame(
  category = rep(c("A", "B", "C", "D"), each = 2),
  group = rep(c("Group 1", "Group 2"), times = 4),
  value = c(2, 3, 8, 6, 3, 4, 7, 5)
)

spec_grouped <- list(
  plot = list(
    list(
      mark = "barY",
      data = list(from = "grouped_data"),
      x = "category",
      y = "value",
      fill = list(column = "group")
    )
  )
)

mosaic(spec_grouped, grouped_data = grouped_data)
```

## Use Cases

- **Categorical Data**: Perfect for displaying values across discrete categories
- **Comparisons**: Easy to compare values between different categories
- **Frequency Counts**: Ideal for showing counts or frequencies of categorical data
- **Survey Results**: Commonly used for displaying survey response distributions