---
title: "lineY Mark"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mosaicr)
```

The `lineY` mark creates line charts, connecting data points with lines. It's ideal for showing trends over continuous data.

## Basic Example

```{r}
# Sample data
data <- data.frame(
  category = c("A", "B", "C", "D", "E", "F", "G", "H"),
  value = c(2, 8, 3, 7, 5, 4, 6, 1)
)

spec <- list(
  plot = list(
    list(
      mark = "lineY",
      data = list(from = "sample_data"),
      x = "category",
      y = "value",
      stroke = "steelblue"
    )
  )
)

mosaic(spec, sample_data = data)
```

## Customization Options

### With Markers

```{r}
spec_markers <- list(
  plot = list(
    list(
      mark = "lineY",
      data = list(from = "sample_data"),
      x = "category",
      y = "value",
      stroke = "steelblue",
      marker = "circle"
    )
  )
)

mosaic(spec_markers, sample_data = data)
```

### Curved Lines

```{r}
spec_curved <- list(
  plot = list(
    list(
      mark = "lineY",
      data = list(from = "sample_data"),
      x = "category",
      y = "value",
      stroke = "steelblue",
      curve = "monotone-x",
      marker = "circle"
    )
  )
)

mosaic(spec_curved, sample_data = data)
```

### Multiple Lines

```{r}
# Time series data
time_data <- data.frame(
  date = rep(as.Date("2023-01-01") + 0:11, times = 2),
  series = rep(c("Series A", "Series B"), each = 12),
  value = c(
    c(10, 12, 8, 15, 18, 14, 20, 16, 22, 19, 25, 21),
    c(8, 10, 12, 9, 14, 16, 13, 18, 15, 20, 17, 23)
  )
)

spec_multiple <- list(
  plot = list(
    list(
      mark = "lineY",
      data = list(from = "time_data"),
      x = "date",
      y = "value",
      stroke = list(column = "series"),
      marker = "circle"
    )
  )
)

mosaic(spec_multiple, time_data = time_data)
```

## Use Cases

- **Time Series**: Perfect for displaying data changes over time
- **Trends**: Shows patterns and trends in continuous data
- **Comparisons**: Multiple lines can compare different series
- **Scientific Data**: Common in research for showing relationships between variables