---
title: "Get Started"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mosaic)
```

Mosaic allows creating scalable and interactive data visualizations using the 
[Mosaic](https://idl.uw.edu/mosaic/) framework directly from R. You can use
Mosaic to create a wide range of visualizations.

## Installation

To install the Mosaic package, you can use the following command in R:

```R
pak::pkg_install("dfalbel/mosaic")
```

## Basic Usage

To create mosaic plots, use the `mosaic` function. First define a specification. The spec is a list that describes the plot(s) you want to create with mosaic as well as the data to be used and optionally more attributes. See the [specification documentation](https://idl.uw.edu/mosaic/api/spec/format.html) for more details.

At its simplest, a spec can be a list with a single plot definition. Here is an example of a simple scatter plot using the built-in `palmerpenguins` dataset:

```{r}
spec <- list(
  plot = list(
    list(
      mark = "dot",
      data = list(from = "penguins"),
      x = "body_mass",
      y = "flipper_len"
    )
  )
)
```

Similar to other grammars, a plot consists of marks — graphical primitives such as bars, areas, and lines—which serve as chart layers. Mosaic uses the semantics of Observable Plot, such that each plot has a dedicated set of encoding channels with named scale mappings such as x, y, color, opacity, etc. See [marks API reference](https://idl.uw.edu/mosaic/api/vgplot/marks.html) for a full list of available marks and their channels.

```{r}
mosaic(
  spec,
  penguins = head(penguins, 50)
)
```
